---
title: "Make pictures"
output: html_document
---

```{r}
library(genstats)
library(bigsnpr)
library(bigstatsr)
library(dplyr)
library(ggplot2)
```



# Getting data

```{r eval=FALSE}
data = snp_attach('example_data.rds')
cor = 10
```


# GWAS
```{r eval=FALSE}
# Models
gwas_summary = GWAS(G = data$genotypes, y=data$fam$pheno_0, p=5e-7, ncores=cor)
gwaslog_summary = GWAS(G = data$genotypes, y=data$fam$pheno_0, p=5e-7, logreg = TRUE, ncores=cor)
beta = data$map$beta

# Plots
plt_gwas_man = manhattan_plot(gwas_summary, beta, thresholds = c(5e-7, 0.05))
plt_gwas_scat = scatter_plot(gwas_summary, beta)
plt_gwas_pow = power_plot(gwas_summary, beta)
plt_log_gwas_man = manhattan_plot(gwaslog_summary, beta, thresholds = c(5e-7, 0.05))
plt_log_gwas_pow = power_plot(gwaslog_summary, beta)
```


# LTFH
```{r eval=FALSE}
# Models
est = LTFH(data, n_sib=2)
y_ltfh = est$l_g_est_0
ltfh_summary = GWAS(G = data$genotypes, y=y_ltfh, p=5e-7, ncores=cor)

#Plots
plt_ltfh_man = manhattan_plot(ltfh_summary, beta, thresholds = c(5e-7, 0.05))
plt_ltfh_scat = scatter_plot(ltfh_summary, beta)
plt_ltfh_pow = power_plot(ltfh_summary, beta)
```

# Compare

## Confussion

```{r eval=FALSE}
true_causal = (data$map$beta != 0) - 0
estimate <- tibble(estimate_gwas = gwas_summary$estim, estimate_gwasLTFH = ltfh_summary$estim)
causal <- tibble(causal_gwas = gwas_summary$causal_estimate, 
                 causal_gwaslog = gwaslog_summary$causal_estimate,
                 causal_gwasLTFH = gwas_LTFH$causal_estimate)

df = causal %>% 
  pivot_longer(cols = everything(), names_to = "method") %>% 
  group_by(method) %>% 
  summarise("11" = sum(value == 1 & true_causal == 1),                                                
            "10" = sum(value == 1 & true_causal == 0),                                                
            "01" = sum(value == 0 & true_causal == 1),                                                
            "00" = sum(value == 0 & true_causal == 0)) %>% 
  pivot_longer(cols = !method) %>% group_by(method) %>% 
  mutate(
    'Prediction' = as.character(c(1,1,0,0)),
    'Actual' =  as.character(c(1, 0, 1, 0)))


conf_compare = df %>% 
  ggplot(aes(x = Actual, y = Prediction)) +
  geom_tile(aes(fill = value), colour = "white", show.legend = FALSE) +
  geom_text(ggplot2::aes(label = sprintf("%1.0f", value)), vjust = 1) +
  scale_fill_gradient(high = "firebrick", low = 'dodgerblue3', trans='pseudo_log') +
  facet_wrap(~method)


```


## Powerplot

```{r eval=FALSE}

gwas_df = gwas_summary %>% 
  mutate(true_causal = (beta != 0) - 0) %>%
  filter(true_causal == 1) %>%
  arrange(abs(estim)) %>%
  mutate(power = cumsum(causal_estimate)/sum(true_causal), Method='GWAS') %>%
  select(estim, power, Method)

gwaslog_df = gwaslog_summary %>% 
  mutate(true_causal = (beta != 0) - 0) %>%
  filter(true_causal == 1) %>%
  arrange(abs(estim)) %>%
  mutate(power = cumsum(causal_estimate)/sum(true_causal), Method='GWASlog') %>%
  select(estim, power, Method)

ltfh_df = ltfh_summary %>% 
  mutate(true_causal = (beta != 0) - 0) %>%
  filter(true_causal == 1) %>%
  arrange(abs(estim)) %>%
  mutate(power = cumsum(causal_estimate)/sum(true_causal), Method='LT-FH') %>%
  select(estim, power, Method)


pow_compare = bind_rows(gwas_df, gwaslog_df, ltfh_df) %>%
  ggplot() + 
  geom_line(aes(x = estim, y = power, color=Method)) + xlim(-0.3, 0.3)

```

## Manhattan

```{r eval=FALSE}

lgnd_m = cowplot::get_legend(plt_gwas_man + theme(legend.position = "top"))

plt_gwas_man_m = plt_gwas_man + theme(legend.position="none")
plt_log_gwas_man_m = plt_log_gwas_man + theme(legend.position="none")
plt_ltfh_man_m = plt_ltfh_man + theme(legend.position="none")

man_compare = gridExtra::grid.arrange(plt_gwas_man_m, 
                                      plt_log_gwas_man_m, 
                                      plt_ltfh_man_m,
                                      lgnd_m, 
                                      ncol=3,
                                      nrow=2,
                                      layout_matrix = rbind(c(1,2, 3), c(4,4,4)),
                                      widths = c(4, 4, 4),  
                                      heights = c(5, 0.5),
                                      top = c("Manhattan plots for GWAS, GWAS using logistic regression, and LT-FH"))




lgnd_s = cowplot::get_legend(plt_gwas_scat + theme(legend.position = "top"))
plt_gwas_scat_s <- plt_gwas_scat + theme(legend.position="none")
plt_ltfh_scat_s <- plt_ltfh_scat + theme(legend.position="none")

scat_compare = gridExtra::grid.arrange(plt_gwas_scat_s, 
                                      plt_ltfh_scat_s,
                                      lgnd_m, 
                                      ncol=2,
                                      nrow=2,
                                      layout_matrix = rbind(c(1,2), c(3,3)),
                                      widths = c(4, 4),  
                                      heights = c(5, 0.5),
                                      top = c("Scatter plots for GWAS and LT-FH"))

```


# Prediction

## getting data
```{r eval=FALSE}
setwd('E:/Data')
train_set = snp_attach("example_data_small.rds")
test_set <- snp_attach("example_data_small_test.rds")
```

## Parameters


```{r eval=FALSE}
est_p = LTFH(train_set, n_sib=2)
y_ltfh_p = est_p$l_g_est_0

prs <- PRS_cross(train_data = train_set, y = y_ltfh_p, cross_folds = 4)

prs_auc = prs_plot(PRS = prs, train_set, 'AUC')
prs_mse = prs_plot(PRS = prs, train_set, 'MSE')
prs_r2 = prs_plot(PRS = prs, train_set, 'R2')


eval_prs_auc = gridExtra::grid.arrange(prs_auc$fold, prs_auc$mean, ncol=2)
eval_prs_mse = gridExtra::grid.arrange(prs_mse$fold, prs_mse$mean, ncol=2)
eval_prs_r2 = gridExtra::grid.arrange(prs_r2$fold, prs_r2$mean, ncol=2)

```



```{r eval=FALSE}
dc_plt = decision_cross(train_set, y_ltfh_p, cross_folds = 4, thr = 3, bounds=seq(0.2, 0.3, 0.01))

```

## LTFH
```{r eval=FALSE}
# Models
ltfh_summary_p = GWAS(G = train_set$genotypes, y=y_ltfh_p, p=5e-7, ncores=cor)
prs1 <- bigsnpr::snp_PRS(G = train_set$genotypes, betas.keep =ltfh_summary_p$estim, lpS.keep = -log10(ltfh_summary_p$p_vals), thr.list = 3)
m = pred_model(train_set, y_ltfh_p, 3, LogReg = FALSE, ncores=cor)
pred = prediction(test_set, m, 3)

# Plots
prs_model = train_set$fam %>%
          ggplot2::ggplot(ggplot2::aes(x=prs1, y=y_ltfh_p)) +
          ggplot2::geom_point(aes(color=as.character(pheno_0)), alpha=0.4) +
          ggplot2::xlab('PRS') +
          ggplot2::ylab('Estimated genetic liability or phenotype') +
          ggplot2::labs(color='Phenotype')

prs_lt_train = train_set$fam %>% ggplot(aes(x=prs1, y=l_g_0)) + geom_point(aes(color=(as.character(pheno_0))), alpha=0.4) + 
  xlab('PRS from train set') + 
  ylab('True genetic liability') + 
  labs(color='Phenotype') 

prs_lt_test = test_set$fam %>% ggplot(aes(x=pred, y=l_g_0)) + 
  geom_point(aes(color=(as.character(pheno_0))), alpha=0.4)  + 
  xlab('PRS from test set') + 
  ylab('True genetic liability') + 
  labs(color='Phenotype')
```



## Result
```{r eval=FALSE}
preds = (pred > 0.28)-0
confusion_matrix = as_tibble(table('Predicted' = preds[,1], 'Actual'= test_set$fam$pheno_0))

dc_result = confusion_matrix %>% 
  ggplot(aes(x = Predicted, y = Actual)) +
  geom_tile(aes(fill = n), show.legend = FALSE) +
  geom_text(aes(label = sprintf("%1.0f", n)), vjust = 1) +
  scale_fill_gradient(high = "firebrick", low = 'dodgerblue3', trans='pseudo_log')
```



## GWAS
Models
```{r eval=FALSE}
gwas_summary_p = GWAS(G = train_set$genotypes, y=train_set$fam$pheno_0, p=5e-7, ncores=cor)
gwaslog_summary_p = GWAS(G = train_set$genotypes, y=train_set$fam$pheno_0, p=5e-7, logreg = TRUE, ncores=cor)
y01 = train_set$fam$pheno_0

# Pheno + linear
prs2 = bigsnpr::snp_PRS(G = train_set$genotypes, betas.keep =gwas_summary_p$estim, lpS.keep = -log10(gwas_summary_p$p_vals), thr.list = 3)
m1 = pred_model(train_set, y01, 3, LogReg = FALSE, ncores=cor)
pred1 = prediction(test_set, m1, 3)

# Pheno + logistic
prs3 = bigsnpr::snp_PRS(G = train_set$genotypes, betas.keep =gwaslog_summary_p$estim, lpS.keep = -log10(gwaslog_summary_p$p_vals), thr.list = 3)
m2 = pred_model(train_set, y01, 3, LogReg = TRUE, ncores=cor)
pred2 = prediction(test_set, m2, 3)

```


Creating plot
```{r eval=FALSE}
# Pheno + linear
prs_lin01_train = train_set$fam %>% ggplot(aes(x=prs2, y=l_g_0)) + 
  geom_point(aes(color=(as.character(pheno_0))), alpha=0.4) + 
  xlab('PRS from train set using linear regression') + 
  ylab('True genetic liability') + 
  labs(color='Phenotype') + theme(legend.position = "top")
  
prs_lin01_test = test_set$fam %>% ggplot(aes(x=pred1, y=l_g_0)) + 
  geom_point(aes(color=(as.character(pheno_0))), alpha=0.4,  show.legend = FALSE)  + 
  xlab('PRS from test set using linear regression') + 
  ylab('True genetic liability') + 
  labs(color='Phenotype')

# Pheno + logistic
prs_log01_train = train_set$fam %>% ggplot(aes(x=prs3, y=l_g_0)) + 
  geom_point(aes(color=(as.character(pheno_0))), alpha=0.4, show.legend = FALSE) + 
  xlab('PRS from train set using logistic regression') + 
  ylab('True genetic liability') + 
  labs(color='Phenotype') 

prs_log01_test = test_set$fam %>% ggplot(aes(x=pred2, y=l_g_0)) + 
  geom_point(aes(color=(as.character(pheno_0))), alpha=0.4, show.legend = FALSE)  + 
  xlab('PRS from test set using logistic regression') + 
  ylab('True genetic liability') + 
  labs(color='Phenotype')

# Grid
lgnd = cowplot::get_legend(prs_lin01_train)
prs_lin01_train <- prs_lin01_train + theme(legend.position="none")

prs_pheno = gridExtra::grid.arrange(prs_lin01_train, 
                        prs_lin01_test, 
                        prs_log01_test,
                        prs_log01_test,
                        lgnd, 
                        ncol=2, 
                        nrow=3, 
                        layout_matrix = rbind(c(1,2), c(3,4), c(5,5)),
                        widths = c(4, 4), 
                        heights = c(5, 5, 0.5))


```

# Saving plots

```{r eval=FALSE}
plt_gwas_man 
plt_gwas_scat
plt_gwas_pow 
plt_log_gwas_man 
plt_log_gwas_pow
plt_ltfh_man 
plt_ltfh_scat
plt_ltfh_pow
eval_prs_auc
eval_prs_mse
eval_prs_r2
prs_model
dc_plt
prs_lt_train
prs_lt_test
dc_result
pow_compare
man_compare
scat_compare
conf_compare
```

Filenames(jpg):
'GWAS_manhattan'
'GWAS_scatter'
'GWAS_power'

'GWASlog_manhattan'
'GWASlog_power'

'pow_compare'
'man_compare'
'scat_compare'
'conf_compare'

'LTFH_manhattan'
'LTFH_scatter'
'LTFH_power'


'prs_true_test'
'prs_true_train'
'prs_est_train'
'prs_pheno'

'decision'
'result_decision'

'eval_prs_auc'
'eval_prs_mse'
'eval_prs_r2' 

```{r eval=FALSE}

## GWAS
#ggsave('GWAS_manhattan.jpg', plt_gwas_man, height = 20, width = 20, units='cm')
#ggsave('GWAS_scatter.jpg', plt_gwas_scat, height = 20, width = 20, units='cm')
#ggsave('GWAS_power.jpg', plt_gwas_pow, height = 20, width = 20, units='cm')

## GWAS log
#ggsave('GWASlog_manhattan.jpg', plt_log_gwas_man, height = 20, width = 20, units='cm')
#ggsave('GWASlog_power.jpg', plt_log_gwas_pow, height = 20, width = 20, units='cm')

## Compare
#ggsave('pow_compare.jpg', pow_compare, height = 20, width = 20, units='cm')
#ggsave('man_compare.jpg', man_compare, height = 20, width = 20, units='cm')
#ggsave('scat_compare.jpg', scat_compare, height = 20, width = 20, units='cm')
#ggsave("confusion_comparison.jpg", conf_compare, width = 20, height = 20, units = "cm")

## LT-FH
#ggsave('LTFH_manhattan.jpg', plt_ltfh_man, height = 20, width = 20, units='cm') 
#ggsave('LTFH_scatter.jpg', plt_ltfh_scat, height = 20, width = 20, units='cm')
#ggsave('LTFH_power.jpg', plt_ltfh_pow, height = 20, width = 20, units='cm')

## Prediction 
#ggsave('prs_est_train.jpg', prs_model, height = 20, width = 20, units='cm')
#ggsave('prs_true_train.jpg', prs_lt_train, height = 20, width = 20, units='cm')
#ggsave('prs_true_test.jpg', prs_lt_test, height = 20, width = 20, units='cm')
#ggsave('prs_pheno.jpg', prs_pheno, height = 20, width = 20, units='cm')
#ggsave('result_decision.jpg', dc_result, height = 20, width = 20, units='cm')

## PRS pram    
#ggsave('decision.jpg', dc_plt, height = 20, width = 20, units='cm')
#ggsave('eval_prs_auc.jpg', eval_prs_auc, height = 20, width = 20, units='cm')
#ggsave('eval_prs_mse.jpg', eval_prs_mse, height = 20, width = 20, units='cm')
#ggsave('eval_prs_r2.jpg', eval_prs_r2, height = 20, width = 20, units='cm')


``` 
 



